surfing

### 2021-07-20 inst/surfing.Rmd
analyse using just the pairs and compare with whole dataset

```{r definesurfing, cache=TRUE}
remove <- c(wildcards, QS, "Vries")
wanted <- which(!(rownames(surfing_table) %in% remove)) # used at directlogplotofstrengths
surfing <- hyper2()
for(v in o){
   r <- v[-1]  # first entry is venue
   numberretained <- sum(!(r %in% remove))
   if(numberretained > 1){
     r <- r[!(r %in% remove)]
     surfing <- surfing + race(r)
    }
}	
summary(surfing)
```

@@

@@

```{r consistencycheck,cache=TRUE}
consistency(surfing)
```

@@

```{r showsurfingmaxppie, echo=FALSE, fig.height=8,fig.cap="dot chart of result"}
dotchart(surfing_maxp)
dotchart(log10(surfing_maxp))
pie(surfing_maxp)
```

@@

```{r calculate_compatriot_monster_LF,cache=FALSE}
scomp <- hyper2()
scomp3 <- hyper2()  # scomp3 is the likelihood function for just the races with 3 competitors, 2 of one nationality and one odd-man-out.
for(v in o){
   r <- v[-1]  # first entry is venue; "r" for "race"
   numberretained <- sum(!(r %in% remove))
   if(numberretained > 1){
     r <- r[!(r %in% remove)]
     if(numberretained ==3){
       nationalities <- c(nationality(r[1]),nationality(r[2]),nationality(r[3]))
       if(length(unique(nationalities)) == 2){
         if(sum(nationalities %in% nationality(r[1])) == 1){ # winner is the odd one out
           Hjj <- pwa(race(r),pwa=r[1])
           print("odd one out is number 1")
           print(100*surfing_maxp[r])
         } else if(sum(nationalities %in% nationality(r[2]) ) == 1){ # second best is the odd one out
           Hjj <- pwa(race(r),pwa=r[2])
           print("odd one out is number 2")
           print(100*surfing_maxp[r])
         } else if(sum(nationalities %in% nationality(r[3])) == 1){  # third best is the odd one out
	   Hjj <- pwa(race(r),pwa=r[3])
           print("odd one out is number 3")
           print(100*surfing_maxp[r])
         } else {
           stop("this cannot happen")
         }
         print(Hjj)
         print("---")
         scomp <- scomp + Hjj      # add compatriot-monster modified likelihood 
         scomp3 <- scomp3 + Hjj
       } else {                    # 3 in race, but no compatriot monster...
         scomp <- scomp + race(r)  # ... so use regular Plackett-Luce for 3-competitor race
       }
     } else {                    # if(numberretained==3) closes...
       scomp <- scomp + race(r)  # numberretained !=3; use regular Plackett-Luce
     }
  } # if(numberretained>1) closes
} # for(v in o) loop closes
```

```{r maximize_compatriot_monster,cache=FALSE}
maxs <- maxp(scomp)
maxs
```

>>>>

```{r definesurfing, cache=TRUE}
remove <- c(wildcards, QS, "Vries")
wanted <- which(!(rownames(surfing_table) %in% remove)) # used at directlogplotofstrengths
surfing <- hyper2()
surfing_pairs <- hyper2()
for(v in o){
   r <- v[-1]  # first entry is venue
   numberretained <- sum(!(r %in% remove))
   if(numberretained > 1){
     r <- r[!(r %in% remove)]
     surfing <- surfing + race(r)
     if(numberretained == 2){
     r <- r[!(r %in% remove)]
     surfing_pairs <- surfing_pairs + race(r)
     }	      
    }
}	
summary(surfing)
summary(surfing_pairs)
```

```{r showsurfing_pairs}
summary(surfing_pairs)
```

@@

Now compare `surfing` with `surfing_pairs`.  First calculate

```{r maxpcalcpairs,cache=TRUE}
surfing_pairs_maxp <- maxp(surfing_pairs)
```

then plot

```{r plotmaxppairs}
surfing_pairs_maxp
surfing_maxp
par(pty='s')
plot(surfing_pairs_maxp,surfing_maxp,asp=1,pch=16)
abline(0,1)
plot(surfing_pairs_maxp,surfing_maxp,asp=1,pch=16,log='xy',xlim=c(0.01,0.2),ylim=c(0.01,0.2))
abline(0,1)
plot(log(surfing_pairs_maxp/surfing_maxp),pch=16)
abline(0,0)
```

@@

```{r consistencycheck,cache=TRUE}
consistency(surfing)
consistency(surfing_pairs)
```

```{r findhessiansurfing,cache=TRUE}
e_all <- eigen(hessian(surfing,indep(surfing_maxp)),T,T)$values
e_pairs <- eigen(hessian(surfing_pairs,indep(surfing_pairs_maxp)),T,T)$values
```

```{r showhessiansurfing}
e_all
e_pairs
f <- function(x){prod(sort(x)[-1])}
f(e_all)
f(e_pairs)
f(e_all)/f(e_pairs)
```

@@

```{r echo=FALSE}
`mydotchart` <- 
function (x, labels = NULL, groups = NULL, gdata = NULL, offset = 1/8, 
    ann = par("ann"), xaxt = par("xaxt"), frame.plot = TRUE, 
    log = "", cex = par("cex"), pt.cex = cex, pch = 21, gpch = 21, 
    bg = par("bg"), color = par("fg"), gcolor = par("fg"), lcolor = "gray", 
    xlim = range(x[is.finite(x)]), main = NULL, xlab = NULL, 
    ylab = NULL, ...) 
{
    opar <- par("mai", "mar", "mgp", "cex", "yaxs")
    on.exit(par(opar))
    par(cex = cex, yaxs = "i")
    if (!is.numeric(x)) 
        stop("'x' must be a numeric vector or matrix")
    n <- length(x)
    if (is.matrix(x)) {
        if (is.null(labels)) 
            labels <- rownames(x)
        if (is.null(labels)) 
            labels <- as.character(seq_len(nrow(x)))
        labels <- rep_len(labels, n)
        if (is.null(groups)) 
            groups <- col(x, as.factor = TRUE)
        glabels <- levels(groups)
    }
    else {
        if (is.null(labels)) 
            labels <- names(x)
        glabels <- if (!is.null(groups)) 
            levels(groups)
        if (!is.vector(x)) {
            warning("'x' is neither a vector nor a matrix: using as.numeric(x)")
            x <- as.numeric(x)
        }
    }
    linch <- if (!is.null(labels)) 
        max(strwidth(labels, "inch"), na.rm = TRUE)
    else 0
    if (is.null(glabels)) {
        ginch <- 0
        goffset <- 0
    }
    else {
        ginch <- max(strwidth(glabels, "inch"), na.rm = TRUE)
        goffset <- offset
    }
    nmai <- opar[["mai"]]
    if (ann) 
        nm.2 <- nmai[2L]
    if (!(is.null(labels) && is.null(glabels))) {
        yi <- if (is.null(ylab) || !ann) 
            0
        else offset
        nm.2 <- nmai[4L] + max(yi + linch + goffset, ginch) + 
            1/16
        if (nmai[2L] < nm.2) {
            nmai[2L] <- nm.2
            par(mai = nmai)
        }
    }
    if (is.null(groups)) {
        o <- seq_len(n)
        y <- o
        ylim <- c(0, n + 1)
    }
    else {
        o <- sort.list(as.numeric(groups), decreasing = TRUE)
        x <- x[o]
        groups <- groups[o]
        color <- rep_len(color, length(groups))[o]
        lcolor <- rep_len(lcolor, length(groups))[o]
        pch <- rep_len(pch, length(groups))[o]
        of.1 <- cumsum(c(0, diff(as.numeric(groups)) != 0))
        y <- seq_len(n) + 2 * of.1
        ylim <- range(0, y + 2)
    }
#    plot.window(xlim = xlim, ylim = ylim, log = log)
    lheight <- par("csi")
    if (!is.null(labels)) {
        loffset <- (linch + 0.1)/lheight
        mtext(labels[o], side = 2, line = loffset, at = y, adj = 0, 
            col = color, las = 2, cex = cex, ...)
    }
    abline(h = y, lty = "dotted", col = lcolor)
    points(x, y, pch = pch, col = color, bg = bg, cex = pt.cex/cex)
    if (!is.null(groups)) {
        gpos <- rev(cumsum(rev(tapply(groups, groups, length)) + 
            2) - 1)
        ginch <- max(strwidth(glabels, "inch"), na.rm = TRUE)
        goffset <- (max(linch + offset, ginch, na.rm = TRUE) + 
            1/16)/lheight
        mtext(glabels, side = 2, line = goffset, at = gpos, adj = 0, 
            col = gcolor, las = 2, cex = cex, ...)
        if (!is.null(gdata)) {
            abline(h = gpos, lty = "dotted")
            points(gdata, gpos, pch = gpch, col = gcolor, bg = bg, 
                cex = pt.cex/cex, ...)
        }
    }
    axis(1, xaxt = xaxt)
    if (frame.plot) 
        box()
    if (ann) {
        title(main = main, xlab = xlab, ...)
        mgp <- par("mgp")
        par(mgp = c(max(mgp[1], nm.2/lheight - 1.5), mgp[-1]))
        title(ylab = ylab, ...)
    }
    invisible()
}
```

```{r showsurfingmaxppie, echo=FALSE, fig.height=8,fig.cap="dot chart of result"}
dotchart(surfing_maxp)
dotchart(log10(surfing_maxp))
pie(surfing_maxp)
jj <- log(surfing_maxp)
kk <- log(surfing_pairs_maxp)
names(jj) <- NULL
names(kk) <- NULL
l <- c(-4.6, -1.5)
dotchart(log(surfing_maxp),pch=NA,xlim=l)
mydotchart(jj,col="red" ,pch=16,xlim=l,xaxt='n',frame.plot=FALSE)
mydotchart(kk,col="blue",pch=16,xlim=l,xaxt='n',frame.plot=FALSE)
```

@@

```{r calculate_compatriot_monster_LF,cache=TRUE}
scomp <- hyper2()
scomp3 <- hyper2()  # scomp3 is the likelihood function for just the races with 3 competitors, 2 of one nationality and one odd-man-out.
for(v in o){
   r <- v[-1]  # first entry is venue; "r" for "race"
   numberretained <- sum(!(r %in% remove))
   if(numberretained > 1){
     r <- r[!(r %in% remove)]
     if(numberretained ==3){
       nationalities <- c(nationality(r[1]),nationality(r[2]),nationality(r[3]))
       if(length(unique(nationalities)) == 2){
         if(sum(nationalities %in% nationality(r[1])) == 1){ # winner is the odd one out
           Hjj <- pwa(race(r),pwa=r[1])
           print("odd one out is number 1")
           print(100*surfing_maxp[r])
         } else if(sum(nationalities %in% nationality(r[2]) ) == 1){ # second best is the odd one out
           Hjj <- pwa(race(r),pwa=r[2])
           print("odd one out is number 2")
           print(100*surfing_maxp[r])
         } else if(sum(nationalities %in% nationality(r[3])) == 1){  # third best is the odd one out
	   Hjj <- pwa(race(r),pwa=r[3])
           print("odd one out is number 3")
           print(100*surfing_maxp[r])
         } else {
           stop("this cannot happen")
         }
         print(Hjj)
         print("---")
         scomp <- scomp + Hjj      # add compatriot-monster modified likelihood 
         scomp3 <- scomp3 + Hjj
       } else {                    # 3 in race, but no compatriot monster...
         scomp <- scomp + race(r)  # ... so use regular Plackett-Luce for 3-competitor race
       }
     } else {                    # if(numberretained==3) closes...
       scomp <- scomp + race(r)  # numberretained !=3; use regular Plackett-Luce
     }
  } # if(numberretained>1) closes
} # for(v in o) loop closes
```

```{r maximize_compatriot_monster,cache=TRUE}
maxs <- maxp(scomp)
maxs
```
